---
# YAML metadata
title: "R for Data Science - Manual"
output:
  pdf_document: default
  html_document: 
  # Add table of contents (TOC):
    toc: true
    number_sections: true
    theme: journal
    toc_depth: 3
    # TOC floating to the left of the main document content
    toc_float: 
      collapsed: false
      smooth_scroll: false
  word_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
library(tidyverse)
# add parameter 'echo = FALSE' to the code chunk to prevent printing of the R code that generated the plot. 
# more details: http://rmarkdown.rstudio.com

```

# Basic information

## Useful resources

[!!! The most useful book for R markdown - "R Markdown: The Definitive Guide"](https://bookdown.org/yihui/rmarkdown/)
[Ebook "R for Graduate Students"](https://bookdown.org/yih_huynh/Guide-to-R-Book/)  
[Ebook "R for Data Science"](https://r4ds.had.co.nz/)  
[Ebook "Statistical Inference via Data Science: A ModernDive into R and Tidyverse"](https://moderndive.com/)  
[Gradient generator](https://colordesigner.io/gradient-generator)  
[ggthemes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)  
[R studio cheatsheets](https://www.rstudio.com/resources/cheatsheets/)  
[Introduction to ggplot2](http://r-statistics.co/Complete-Ggplot2-Tutorial-Part1-With-R-Code.html)  
[Themes for Bootstrap](https://bootswatch.com/3/)




## Versions and plugins

To update R version: go to R GUI (old), run 1) install.packages("installr"); library(installr); choose installr option on the tab next to the option "Help", then update... voila!

## Packages installation and loading: 
``` {r, eval = FALSE}

getRversion()

install.packages(c("afex", "tidyverse", "emmeans", "writexl", "readxl", "ggthemes", "ggplot2", "extrafont", "reshape", "reshape2", "gcookbook", "readr", "swirl", "ggpubr", "lubridate", "rmarkdown", "formatR", "knitr", "magrittr", 'glue', 'viridis', 'ggsci', 'gcookbook'))

install.packages('plotly')

```

```{r, results ='hide', message=FALSE}

library(afex); library(emmeans); library(tidyverse); library(writexl); library(readxl); library(ggthemes); library(extrafont); library(reshape); library(reshape2); library(gcookbook); library(readr); library(swirl); library(ggpubr); library(ggthemes); library(lubridate); library(rmarkdown); library(formatR); library(knitr); library(magrittr); library(glue); library(viridis); library(ggsci); library(gcookbook); library(plotly)

# font_import() # for extrafont package
# loadfonts(device = "win") # loading fonts from extrafont

```

## Shortcuts

`Ctrl + Alt + I` : insert an R code chunk
`Ctrl + Shift + Enter` : run the entire code chunk

# Variables

```{r}

var1 = "string_value"

```



# Data types

Check the data type of a variable: `class(variable)`

## Character (str)

**Methods**:
`length()`

```{r}
# STRING
var1 = "string"
class(var1)

# f-string
a = 5
b = glue('Value: {a}')
print(b)

```

## Integer

```{r}

```


## Numeric


## Logical

```{r}
a = TRUE
class(a)
```


## Vector

Composed of elements of the same type
Slicing is 1-based, so the the starting index is 1 instead of the usual 0. 

Methods:
- `length(listname)`

```{r}

listNames <- c('a', 'b', 'c', 'd', 'e', 'f')
listNames[1]
listNames[1:3] # Slice the list
listNames[c(1, 3, 6)] # Choose elements at indices 1, 3, and 6


```

## List

Composed of elements of different types. 

```{r}
listNames <- list('a', 15, c("1", "2", "3"), 'd', 'e', 'f')
class(listNames)
listNames[[3]] # To slice, usually we use double brackets
```

## Matrix

```{r}
m1 = matrix(c(1,2,3,4,5,6), nrow=2, ncol=3, byrow=TRUE)
m1
m1[1,3] # Row 1, column 3

```


## ???
```{r}

# DATASET
# Create brand-new dataset
data <- data.frame(Chr = c("Chr1", "Chr2", "Chr2", "Chr3", "Chr4", "Chr11"), 
                   Type = c("1_Promoter", "1_Promoter", "2_Enhancer", "3_Activator", "1_Promoter", "3_Activator"), 
                   Value = c(1e+9, 5e+10, 6e+9, 9e+10, 8e+10, 5e+10))

data$Type

```

# Operators

**Assignment operators**

```{r}
# Assignment operators
a = 5
a <- 5
class(a)

```

**Math operators**
```{r}
# Math operators
a = 5; b = 10
c = a + b
c
```

**Relational operators**
>, <, ==, != 

**Logical operators**
& (and), | (or)


# Flow statements

## IF

```{r}

a = 1

if (a > 2){
  print("'a' is greater than 2")
} else {
  print("not")
}
```


## FOR

```{r}
list1 = c(1, 2, 3)

for (i in list1){
  print(i)
}

```

## WHILE

```{r}
i = 1

while (i < 5){
  print(i)
  i = i + 1
}
```


# Modules and packages

**Pre-loaded datasets**:
```{r}

data() # check a list of pre-loaded datasets

```


## Dplyr ( =pandas)

This is basically the analogue to Python's Pandas. 

```{r}
library(dplyr)

# Read a local file
#df <- read.csv("Data/diamonds.csv")
# Load a pre-loaded dataset from a package
starwars # from the dplyr package

dim(df) # Print dimensions of the dataset
head(df) # Print the first 5 rows
# Filter
filter(starwars, height > 170, mass > 100)
# Select columns
select(starwars, height, mass)
# Choose a randomly-selected sample
sample_n(starwars, 5) # A numbered sample
sample_frac(starwars, 0.01) # Fraction of the dataset

```
An example of a pipeline of a dataset processing:

```{r}

starwars %>% filter(mass < 80, sex == "male") -> males
males

```




# Functionality

* Ctrl + shift + R : creates headers for your R script   
* Alt + shift + K : displays all programmed keyboard shortcuts for R   
* Ctrl + shift + M : write "%>%" (tidyverse pipe)    


## Basic commands

* Save as .csv: ```write_csv(name, "filename.csv")```
* Import data from .txt file: ```read.csv("filename.txt", sep="\t", header=FALSE)```
* read.csv("filename.csv") # import data from csv file
* ironmansuits <- read.csv("C:/Users/evgen/Desktop/R_work_folder/Iron_Man_Suits.csv", stringsAsFactors=TRUE)
* mutate() # creates new columns or modifies current variables in the dataset
* diamonds %>% mutate(depth_times_price = depth * price)
* diamonds %>% mutate(depth_times_price = depth * price, is_cut_ideal = cut == "Ideal")
* analog of f-string in R: library(glue): ```glue("Variable {varname}")```
* Export graph as .tiff: ```ggsave("test.tiff", units="in", width=8, height=6, dpi=300, compression = 'lzw')```


**Operators AND (&), OR (|)**  

* diamonds %>% filter(cut == "Ideal", carat == "0.23", price > 400)
* select() # gives us specific columns
* diamonds %>% select (cut, price)
* arrange() # arranges the values in data frame by values in a variable
* diamonds %>% arrange(price)
* diamonds %>% arrange(desc(price))
* diamonds %>% arrange(price, carat)
* group_by() / summarise()
* diamonds %>% group_by(carat) %>% summarise(mean(price), count = n())
* diamonds %>% group_by(carat) %>% mutate(mean_price = mean(price))
* runif() # generates random deviates of uniform distribution

**Describing data**  

* str # describe the structure of dataset by type of variable
* mean, median, max, min, sd
* dim(my_data) ## show dimensions of a table
* dim(my_data)
* summary(my_data)
* summary(my_data$chrom)
* ?ggplot
* seq(1,10)
* data() # show a list of built-in datasets



## Datasets



```{r, results='hide'}
# Create brand-new dataset
data <- data.frame(Chr = c("Chr1", "Chr2", "Chr2", "Chr3", "Chr4", "Chr11"), 
                   Type = c("1_Promoter", "1_Promoter", "2_Enhancer", "3_Activator", "1_Promoter", "3_Activator"), 
                   Value = c(1e+9, 5e+10, 6e+9, 9e+10, 8e+10, 5e+10))
data$Chr = gsub("Chr", "", data$Chr) # remove "Chr" from column 1
data$Chr <- factor(data$Chr, levels=c(seq(1,11), "X", "Y")) # reorder the chromosomes numerically


# Create dataset from existing
diamonds2 <- tibble(diamonds$cut, diamonds$clarity)
names(diamonds2)[1:4] <- c("Cut", "Clarity") # renames the first 4 columns

# Add new columns
diamonds2 <- diamonds %>% 
  mutate(price200 = price - 200, 
         pricepercarat = price / carat)

# Update the in-table columns with new values
diamonds %>% mutate(cut = recode(cut, "Ideal" = "IDEAL", "Good" = "GOOD"))
```

## Filtering

- ```diamonds %>% filter(cut=="Fair" | cut=="Good", price <= 600)``` 
- ```recode () # modifies the values within a variable```
- ```filter () # gives us specific rows```
- ```population2 <- population %>% filter(country == 'Brazil' | country == 'Argentina', year >= 2005)```
- `df$col1 == "name 1"`
- `df$col1 %in% c("name1", "name2")`

## Data processing

- Rename df columns by name: ```df.rename(columns={"A": "a", "B": "c"})```
- Rename df columns by index: ```df.rename(index={0: "x", 1: "y", 2: "z"})```


[Information below about tidy data is taken from this link.](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html#:~:text=Tidy%20data%20is%20a%20standard,Every%20column%20is%20a%20variable.)

**Dataset** - a collection of quantitative and qualitative values. Every value belongs to a variable and an observation.    

**Untidy (messy) data** - rows and columns are mixed up with observations and variables. 

**Tidy data** - a standard way of mapping the meaning of a dataset to its structure. Variables are in columns and observations - in rows. Tidy data makes it easy to extract variables. 

**Examples**:

Untidy (messy) data - 6 observations but only 3 rows - 'wide' dataset: 

```{r}

#define data
df <- data.frame("."=c('John', 'Jane', 'Mary'), 
                 Treatment_A=c(16, 3, 2),
                 Treatment_B=c(0, 4, 5), 
                 Treatment_C=c(0, 8, 10)
)
df

```

Tidy data structure - three variables - Person/name, Treatment, and Result - 'long/tall' dataset - made by melting the untidy one:

```{r}

df <- data.frame("Person/name"=c('John', 'Jane', 'Mary', 'John', 'Jane', 'Mary', 'John', 'Jane', 'Mary'), 
                 "Treatment"=c('a', 'a', 'a', 'b', 'b', 'b', 'c', 'c', 'c'), 
                 "Result"=c(16, 3, 2, 0, 4, 5, 0, 8, 10)
)
df

```


# Graphs and Figures

By convention, independent variable is on X axis, and dependent variable is on Y axis. 

## Dotplots

**Dataset 'population'**

```{r}

# Filtering a sample dataset "population"
population2 <- population %>% filter(country == 'Brazil' | country == 'Argentina', year >= 2000)
#population2

# Simple dotplot
ggplot(population2, aes(x=year, y=population, color=country)) +
  geom_point(size=4) + 
  scale_x_continuous(breaks = (seq(min(population2$year), 
                                   max(population2$year), 
                                   by = 3))) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.text = element_text(size=13),
        axis.title = element_text(size=15), 
        panel.background = element_blank(), 
        axis.line = element_line(size = 1)) + 
  scale_color_brewer(palette = "Set1") + 
  labs(title = "Population growth in Brazil and Argentina", 
       x = "Year", 
       y = "Population")

# Faceted dotplot
ggplot(population2, aes(x=year, y=population)) +
  geom_point() + 
  facet_wrap(~country, nrow=2)

# Dotplot with callouts
ggplot(population2, aes(x=year, y=population, color=country, label=population)) +
  geom_text(check_overlap = TRUE) + 
  geom_label()
```


## Scatterplots

```{r}
msleep

# Scatterplot with df msleep
ggplot(msleep, aes(x=bodywt, y=sleep_total, colour=vore)) + 
  geom_point() + 
  xlim(0, 7000) + ylim(0, 21)
```

## Histograms


```{r}

ggplot(diamonds, aes(x = price)) + geom_bar()
ggplot(diamonds, aes(x = price, fill = cut)) + geom_bar()
ggplot(diamonds, aes(x=price, fill=cut)) + geom_histogram(binwidth=500) + xlim(0,10000)

```

## Bar graphs

For geom_bar, aggregate numbers of rows for each x-value can be counted by default with argument stat=count. Alternatively, the values can be provided manually for the bar graph by stat="identity". 

```{r}

ggplot(diamonds, aes(x=cut)) + geom_bar()

ggplot(diamonds, aes(x=cut, y=price)) + geom_bar(stat='identity')

ggplot(diamonds, aes(x=cut, fill=cut)) + 
  geom_bar() + 
  geom_text(stat="count", aes(label=..count..), vjust=1.5, size=3.5, colour='white')

ggplot(diamonds, aes(x=cut, y=price, fill=cut)) + geom_col()


```

## Stacked component bar graphs

```{r}

ggplot(diamonds, aes(x=cut, fill=color)) + 
  geom_bar()

starwars %>% 
  filter(hair_color == "black" |       # | = OR
           hair_color == "brown") %>% 
  drop_na(sex) %>% 
  ggplot(aes(x = hair_color, fill = sex)) + 
  geom_bar(alpha = 0.7) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + 
  labs(title = "Gender and Hair Colour", 
       x = "Hair Colour", 
       y = "Number")

```

## Stacked percentage bar graphs

```{r, warning=FALSE}

library(reshape)

# Example 1
starwars
starwars %>% 
  filter(hair_color == "black" |       # | = OR
           hair_color == "brown") %>% 
  drop_na(sex) %>% 
  ggplot(aes(x = sex, fill = hair_color)) + 
  geom_bar(position = "fill", 
           alpha = 0.5) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + 
  labs(title = "Gender and Hair Colour", 
       x = "Sex", 
       y = "Percentage")

# Example 2
relig_income
relig_income %>% 
  filter(religion == "Atheist" |
         religion == "Mormon" |
         religion == "Buddhist") %>%
  melt(id.vars = "religion") %>% 
  group_by(religion, variable) %>% 
  ggplot(aes(x = religion, y = value, fill = variable)) +
  geom_bar(stat="identity", position = "fill") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.title = element_text()) + 
  scale_color_brewer(palette = "Set1") +
  labs(title = "Distribution of income levels by religion", 
       x = "Religion", 
       y = "Percentage")

# library gcookbook
cabbage_exp
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) +
  geom_col(position = "fill", width=0.5) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + 
    scale_fill_lancet() + # module ggsci - palettes from sci journals
  labs(title = "Title", 
       x = "date", 
       y = "Weight, kg")

# Custom dataset
df <- data.frame (Type_of_epistasis = c('negative', 'positive', 'negative', 'positive'), 
                  n = c(250, 750, 300, 700), 
                  algorithm = c('algorithm1', 'algorithm1', 'algorithm2', 'algorithm2'), 
                  percentage = c(0.25, 0.75, 0.3, 0.7),
                  percentage_100 = c(25, 75, 30, 70))
df
ggplot(df, aes(x=algorithm, y=percentage, fill=Type_of_epistasis)) +
  geom_col(position='fill', colour='black') + 
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(label = percentage_100), size = 4, position=position_stack(vjust=0.5), color='white') +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + 
  scale_fill_lancet() + # module ggsci - palettes from sci journals
  labs(title = "Distributions of epistasis between two algorithms", 
       x = "Algorithm", 
       y = "Distribution of types of epistasis, %")


```

## Grouped bar graphs

```{r}

ggplot(diamonds, aes(x = cut, fill = clarity)) + geom_bar(position="dodge")

# grouped facet
starwars %>% 
  filter(hair_color %in% c("black", "brown")) %>% 
  drop_na(sex) %>% 
  ggplot(aes(sex)) + 
  geom_bar(aes(fill = sex), alpha = 0.5) + 
  facet_wrap( ~ hair_color)

# grouped bar graphs with error bars
head(ToothGrowth)
df <- ToothGrowth
df$dose <- as.factor(df$dose)
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}
df2 <- data_summary(ToothGrowth, varname="len", 
                    groupnames=c("supp", "dose"))

## Convert dose to a factor variable
df2$dose=as.factor(df2$dose)
head(df2)
p<- ggplot(df2, aes(x=dose, y=len, fill=supp)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2,
                position=position_dodge(.9)) 
print(p)
## Finished bar plot
p+labs(title="Tooth length per dose", x="Dose (mg)", y = "Length")+
  theme_classic() +
  scale_fill_manual(values=c('#999999','#E69F00'))

```

## Boxplots

```{r}

ggplot(diamonds, aes(x=cut, y=price)) + geom_boxplot()
ggplot(diamonds, aes(x=cut, y=price, fill=cut)) + geom_boxplot()


```

## Violin plots

```{r}

starwars 

starwars %>% 
  filter (sex %in% c('male', 'female')) %>% 
  ggplot(aes(x=sex, y=mass)) + geom_violin()

starwars %>% 
  filter (sex %in% c('male', 'female')) %>% 
  ggplot(aes(x=sex, y=mass, fill=sex)) + 
  geom_violin(adjust=0.7) + # default adjust is 1. Lower means finer resolution
  guides(fill=FALSE) # make the entries coloured but delete the legend

```


## Density plots

They are like histograms, but with probability instead od count

```{r, warning=FALSE}

starwars %>% 
  filter(gender != 'NA') %>% 
  ggplot(aes(x=height, fill=gender)) + geom_density(alpha=0.5)

starwars %>% 
  filter(gender != 'NA') %>% 
  ggplot(aes(x=height, fill=gender)) + geom_density() + facet_grid( gender ~ .)

starwars %>% 
  filter(gender != 'NA') %>% 
  ggplot(aes(x=height, fill=gender)) + geom_density() + facet_grid( . ~ gender )

starwars %>% 
  filter(gender != 'NA') %>% 
  ggplot(aes(x=height, fill=gender)) + geom_density() + facet_grid( gender ~ sex )


```

## Line graphs

```{r, warning=FALSE}

starwars2 <- starwars %>% filter(sex %in% c('male', 'female'), gender %in% c('masculine', 'feminine')) 


ggplot(starwars2, aes(x=height, y=mass)) + geom_smooth()

ggplot(starwars2, aes(x=height, y=mass, color=sex)) + geom_smooth()

ggplot(starwars2, aes(x=height, y=mass, colour=sex)) + 
  geom_line(size=3) +
  theme(
    axis.line=element_line(size=0.5), 
    panel.background=element_rect(fill=NA, size=rel(20)), 
    panel.grid.minor=element_line(colour=NA), 
    axis.text=element_text(size=16), 
    axis.title=element_text(size=18)) + 
  scale_y_continuous(expand=c(0,0))

Orange
Orange %>% 
  filter(Tree != "1" &
           Tree != "2") %>% 
  ggplot(aes(x = age, y = circumference, colour = Tree)) + 
  geom_point(size = 5, alpha = 0.3) + 
  geom_line(size = 1) + 
  theme_minimal() + 
  labs(title = "Tree age and circumference")


# line graphs with error bars (SEM) + customly-coloured
# SEM, mean, custom colours
library(ggplot2)

sem <- function(x, na.rm = FALSE) {
  out <- sd(x, na.rm = na.rm)/sqrt(length(x))
  return(out)}
pointcolour <- c("Fair" = "yellow", 
                 "Good" = "red", 
                 "Very Good" = "pink", 
                 "Premium" = "blue", 
                 "Ideal" = "black")
#write.csv(diamonds, file="diamonds.csv", row.names=TRUE)


```

## Pie charts

Nothing here right now...

## Tiered faceted graphs

```{r, warning = FALSE, results = 'asis'}

library(ggpubr)

p1 <- ggplot(diamonds[seq(1, 50000, by = 100),], aes(x = carat, y = price, colour = cut)) + geom_point() + facet_wrap( ~ cut, nrow = 1)
p1
p2 <- ggplot(diamonds[seq(1, 50000, by = 100),], aes(y = price, colour = cut)) + geom_boxplot() + facet_wrap( ~ cut, nrow = 1) # can add argument scale = "free_x" to make it automatically scaled
p2
ggarrange(p1, p2, nrow = 2, labels = "AUTO")
# OR the same, just the entire dataset & without creating new variables #
ggarrange(
  # graph 1
  ggplot(diamonds, aes(x = carat, y = price, colour = cut)) + geom_point() + facet_wrap( ~ cut, nrow = 1), 
  # graph 2
  ggplot(diamonds, aes(y = price, colour = cut)) + geom_boxplot() + facet_wrap( ~ cut, nrow = 1),
  # ggarrange settings
  nrow = 2, labels = "AUTO"
)



summary(relig_income)
p1 <- relig_income %>% 
  filter(religion == "Atheist" |
           religion == "Mormon" |
           religion == "Buddhist") %>%
  melt(id.vars = "religion") %>% 
  group_by(religion, variable) %>% 
  ggplot(aes(x = religion, y = value, fill = variable)) +
  geom_bar(stat="identity", position = "fill") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.title.x = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(size = 0.5)) + 
  scale_color_brewer(palette = "Set1") +
  labs(title = "Distribution of income levels by religion", 
       y = "Percentage")
p2 <- relig_income %>% 
  filter(religion == "Atheist" |
           religion == "Mormon" |
           religion == "Buddhist") %>% 
  melt(id.vars = "religion") %>% 
  group_by(religion, variable) %>% 
  ggplot(aes(x = religion, y = value)) + 
  geom_bar(stat="identity") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.title = element_text(), 
        panel.background = element_blank(), 
        axis.line = element_line(size = 0.5)) + 
  scale_color_brewer(palette = "Set1") +
  labs(title = "Total count of observations by religion", 
       x = "Religion", 
       y = "Count")
ggarrange(p1, p2, nrow = 2, labels = "AUTO")

```


# Graph customisation

## Theme settings

* Flip coordinates: + coord_flip()
* Log scale: + scale_y_log10()
* Show fonts: fonts()
* Globally change theme settings: theme_set(theme_gray(base_size=20))


## Colour changes

* Globally reset theme settings to default: theme_set(theme_gray()) 
* Black and white theme: mainplot + theme_bw()
* Manually assign colours: scale_color_manual(values = c("#A6611A", "#DFC27D", "#6e6c6b", "#80CDC1", "#018571"))

## Colour palettes

* library(RColorBrewer)
* display.brewer.all()
* colours() # gives names of different colours
* mainplot + scale_fill_brewer(palette="Set1")
* Setting colour values for each of the 5 variables in my mainplot graph manually: mainplot + scale_fill_manual(values = c("green", "blue", "red", "orange", "grey")) 


## Changing theme

```+ theme()``` argument

* Number size on axis: axis.text = element_text(size = 20) 
* Title size on axis: axis.title = element_text(size = 20)
* Legend: legend.title = element_text(size = 20)
* Background: panel.background = element_rect(fill = "pink")
* Change major gridlines: panel.grid.major = element_line(colour = "blue")
  * size = 0.2
  * NA
* Change minor gridlines: panel.grid.minor = element_line(colour = "red")
* Change tick marks: axis.ticks = element_line(size=2, colour="blue")
  * As options: axis.ticks.x, axis.ticks.y
* Change legend position: legend.position="top"
  * Other arguments: 
  * "bottom" 
  * Bottom left: c(0,0) 
  * Near the top right: c(0.8,0.8)



## Changing labels

```+ labs()``` argument

* title = "Title"
* y = "Title y"
* x = "Title x"





## Publication style

Clear background, axis lines, no box, no grid lines, basic colors, no legend

```{r, eval=FALSE}

# Just for one plot
mainplot + 
  theme(
    axis.line=element_line(size=0.5), 
    panel.background=element_rect(fill=NA, size=rel(20)), 
    panel.grid.minor=element_line(colour=NA), 
    axis.text=element_text(size=16), 
    axis.title=element_text(size=18)) +
  scale_y_continuous(expand=c(0,0)) # to stop the bars from floaing above the x-axis

# Globally apply this theme
theme_set(theme_gray() + 
            theme(
              axis.line = element_line(size=0.5),
              panel.background = element_rect(fill=NA,size=rel(20)), 
              panel.grid.minor = element_line(colour = NA), 
              axis.text = element_text(size=16), 
              axis.title = element_text(size=18)
            ))


```



